### 2023-01-18 paper.Rmd
Converted paper to Rmarkdown for easier output generation.

---
title: "BumpyMatrix: a Bioconductor package to represent a 2-dimensional array of non-scalar objects"
tags:
  - R
  - Bioconductor
  - bioinformatics
authors:
  - name: Aaron T. L. Lun
    orcid: 0000-0002-3564-4813
    affiliation: 1
affiliations:
  - name: Genentech Inc., South San Francisco, USA
    index: 1
date: 17 January 2022
bibliography: ref.bib
---

```{r, echo=FALSE, results="hide"}
library(BiocStyle)
knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE, collapse=TRUE)
set.seed(100)
```

# Summary

# Statement of need

# Basic usage

Let's mock up some data for a multiplexed FISH experiment.
This consists of a `DataFrame` (Bioconductor's wrapper around the base `data.frame` class) where each row contains the coordinates of a single detected transcript molecule in a particular cell.

```{r}
library(S4Vectors)
df <- DataFrame(
    x=rnorm(10000), y=rnorm(10000), 
    gene=paste0("GENE_", sample(100, 10000, replace=TRUE)),
    cell=paste0("CELL_", sample(20, 10000, replace=TRUE))
)
df
```

We use the `splitAsBumpyMatrix()` function to create a `BumpyDataFrameMatrix` where the rows are genes and the columns are cells.
Each entry of this `BumpyDataFrameMatrix` is itself a `DataFrame` that corresponds to a particular gene/cell combination,
and containing the x/y-coordinates for all transcript molecules of that gene in that cell.
Each `DataFrame` may have zero, one or multiple rows, depending on the cell's expression of the gene.

```{r}
library(BumpyMatrix)
mat <- splitAsBumpyMatrix(df[,c("x", "y")], row=df$gene, column=df$cell)
mat
```

We can now treat `mat` in the same manner as any other matrix-like R object.
For example, we can slice by row and column, equivalent to filtering our original `df` by gene or cell respectively.

```{r}
mat[1:5,]

mat[,c("CELL_5", "CELL_10", "CELL_20")]
```

Extracting a row or column of the `BumpyDataFrameMatrix` will return a list of `DataFrame` objects, which is the 1-dimensional analogue to the atomic vectors obtained from extracting a row/column of an ordinary R matrix.
(More specifically, the return value is a Bioconductor `SplitDataFrameList`, which is used as the underlying data representation for the `BumpyDataFrameMatrix` implementation.)
A single `DataFrame` can be extracted from a `BumpyDataFrameMatrix` or a `SplitDataFrameList` with the usual `[[` operator.

```{r}
mat[1,]

mat[1,1][[1]]
```

The `lengths()` function will return the number of entries in each `DataFrame`.
In this context, the return value is a count matrix that quantifies the expression of each gene in each cell.

```{r}
counts <- lengths(mat)
counts[1:10,1:5]
```

# Compatibility with the `SummarizedExperiment`

Of particular interest is the ability to store `mat` inside Bioconductor's `SummarizedExperiment` data structure.
The `SummarizedExperiment` provides a holistic representation of a dataset that contains all experimental assays, feature- and sample-level annotations as well as additional metadata. 
This representation is convenient as it ensures that any operations are synchronised across all aspects of the dataset, avoiding any account-keeping errors across the lifetime of an analysis workflow.
By storing the `BumpyDataFrameMatrix` inside the `SummarizedExperiment`, we can synchronise the manipulations of the transcript coordinates to the rest of the dataset.

```{r}
library(SummarizedExperiment)
se <- SummarizedExperiment(list(positions = mat, counts = lengths(mat)))
se

# Limiting our analysis to the first 10 cells:
sub.se <- se[,1:10]
dim(assay(sub.se, "counts"))

# Our BumpyDataFrameMatrix is also subject to the same operation:
assay(sub.se, "positions")
```

This compatibility also simplifies the consumption of `BumpyDataFrameMatrix` assays by frameworks that operate on a `SummarizedExperiment`.
For example, the **alabaster** framework provides a mechanism for serializing R/Bioconductor objects into language-agnostic file formats, for use with other programming ecosystems like Python or Javascript.
As **alabaster** already provides a mechanism to serialize `SummarizedExperiment` objects to file, it can be easily extended to handle objects containing `BumpyDataFrameMatrix` objects in its assays.

```{r}
library(alabaster.se) # TODO: change to the alabaster umbrella.

# Saving the SummarizedExperiment into a staging directory.
staging <- tempfile()
dir.create(staging)
info <- stageObject(se, staging, "experiment_1")

# Loading the SummarizedExperiment recovers our BumpyDataFrameMatrix.
reloaded <- loadObject(info, staging)
assay(reloaded, "positions")
```

# Other `BumpyMatrix` subclasses

While we have largely focused on the `BumpyDataFrameMatrix`, the _BumpyMatrix_ package also supports other `BumpyMatrix` subclasses.
For example, the `BumpyNumericMatrix` class implements a matrix-like object where each entry is a numeric vector of any length.
To illustrate, we can generate a `BumpyNumericMatrix` instance by extracting the x-coordinates from `mat`.

```{r}
xpos <- mat[,,"x"]
xpos
```

This can be manipulated in the same manner as a `BumpyDataFrameMatrix`, except that each entry is now a numeric vector instead of a `DataFrame`.

```{r}
# Getting the x-coordinates for the first cell:
xpos[,1]

# Extracting the numeric vector directly:
xpos[1,1][[1]]
```

These `BumpyMatrix` classes can also participate in basic arithmetic, logical and mathematical operations, allowing users to perform complex manipulations in an intuitive manner.
Imagine that we have the centroid for each cell, and we wish to compute the distance of each transcript from its cell's centroid.
We can achieve this by applying the usual formula for a 2-dimensional Euclidean distance on our `BumpyNumericMatrix` objects.

```{r}
# Mocking up a centroid x/y coordinate for each cell.
x_cell <- rnorm(ncol(mat))
y_cell <- rnorm(ncol(mat))

# Computing the Euclidean distance of each transcript from the centroid.
# Transpositions are necessary to ensure that the cell centroid coordinates
# are recycled along the cells, not along the features.
squared_x <- (t(mat[,,"x"]) - x_cell)^2 
squared_y <- (t(mat[,,"y"]) - y_cell)^2
dist <- t(sqrt(squared_x + squared_y))
dist
```

Further imagine that we wish to filter `mat` to remove transcripts that are more than, say, 2 units away from the centroid.
This is possible by converting our `BumpyNumericMatrix` into a `BumpyLogicalMatrix` via the usual logical `<` operation.
The resulting `BumpyLogicalMatrix` can then be used to subset our original `BumpyDataFrameMatrix`, which will apply the filter of the former to the rows of each individual `DataFrame` in the latter.

```{r}
keep <- dist <= 2
keep

# We can use a BumpyLogicalMatrix as a index into the BumpyDataFrameMatrix.
filtered <- mat[keep]
filtered
```

# References